Cycles of the logistic map

### 2013-05-01 doc/pc.tex
doc: a simplified version

In an $n$-cycle, $n$ is the smallest positive integer
  that allows $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
Thus, any $x_k$ in an $n$-cycle of $f$ must be a fixed point of $f^n$
%
\big[the converse is, however, untrue, for a fixed point of $f^n$
  can also be a fixed point of $f^d$ of a divisor $d$ of $dn$:
  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big],
%
%
and a cycle can be classified as stable or unstable
  according to the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_1) \cdots f'(x_n) \Big| \le 1,
\label{eq:der}
\end{equation}
The set of the $n$ points within the cycle $x_1. \ldots, x_n$
   is called the \emph{orbit}.
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz}.
%
%
%

@@

To simplify the calculation, we first change variables \cite{saha} by
%
%\begin{equation}
\[
    x^{\mathrm{new}} \equiv r(x^{\mathrm{old}} - 1/2),
    \quad R \equiv r(r-2)/4,
\]
%\label{eq:chvars}
%\end{equation}
%
and rewrite the map, in terms of $x^{\mathrm{new}}$, as
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv R - {x_k}^2.
  %\tag{$\ref{eq:logmap}'$}
\label{eq:logmaps}
\end{equation}
%
%\refeq{logmaps} simplifies \refeq{logmap} as $x_k$ now only occurs once.
%
We will solve the cycle boundaries as zeros of some polynomials of $R$,
and the corresponding polynomials for $r$ can be obtained by $R \rightarrow r(r-2)/4$.
%
%the numerical $r$ values can always be obtained from $r = 1 + \sqrt{1 + 4R}$.
%

@@

%Let us consider the first step of finding the fixed points of $f^n$.
At the first glance, the polynomials of the fixed points of $f^n$
  can be obtained by brute force.
We can solve
  \refeqs{logmaps} and express $x_1, \ldots, x_n$
  in terms of $R$,
  and then plug the solution into \req{der}.
The result would contain $R$ only,
  and is therefore the answer.
But since \refeqs{logmaps} are nonlinear,
  it is nontrivial to reduce the final equation of $R$
  into a polynomial form.
Nonetheless, on a computer, one can
  construct a (polynomial) Gr\"obner basis \cite{kk1}
  that gradually eliminates $x_k$ out of the $n+1$ equations.
As a special case, if $x_2, \dots, x_n$ are eliminated first,
  the final elimination of $x_1$ can be accomplished
  by setting the resultant of $f^n(x_1) = x_1$ and
  \refeq{der} or $\frac{d}{dx}f^n (x_1) = \pm 1$
  to zero \cite{burm},
  which furnishes the needed polynomial equation of $R$.
%
These approaches, however, do not
  exploit the cyclic structure of \refeqs{logmaps},
  and can thus be improved by the following alternative.

@@

A polynomial is cyclic if it is invariant
  under the cycling of variables
  $x_1 \rightarrow x_2, x_2 \rightarrow x_3,
  \ldots, x_n \rightarrow x_1$,
  e.g., $a(\vx) = x_1 x_2 + x_2 x_3 + x_3 x_4 + x_4 x_1$
  and $b(\vx) = x_1 x_3 + x_2 x_4$,
  where $\vx \equiv \{x_1, \ldots, x_n\}$ and $n = 4$.
Note that a cyclic polynomial is not necessarily from a symmetric one,
  which is invariant under the exchange of any two $x_k$ and $x_{j}$ ($j \ne k$);
  e.g., $a(\vx)$ and $b(\vx)$ are not symmetric,
  but $a(\vx) + b(\vx)$ is.

@@

We now consider
  a subset of \emph{square-free} cyclic polynomials,
  which have no square or higher powers of any $x_k$,
  e.g., $C_{12}(\vx) = x_1 x_2 + \cdots$ is square-free,
  $C_{112}(\vx) = {x_1}^2 x_2 + \cdots$ is not,
  see \reftab{sqrfreepoly} for more examples.
The label $p$ of a square-free polynomial
  has no repeated index.
We denote the set of all square-free $p$ by
$\Bset = \{0, 1, 12, 13, \ldots, 123, 124, \ldots, 12\dots n\}$
%  ($n \ge 4$),
such that its size $|\Bset|$ equals the number $\NB$ of
  square-free cyclic polynomials.

@@

We first show that
  the square-free cyclic polynomials
  serve as a basis for expanding cyclic polynomials:

\begin{theorem}
  For the logistic map \refeq{logmaps},
  any cyclic polynomial $K(\vx)$
  formed by the $n$-cycle points
  $\vct x = \{x_1, \ldots, x_n\}$
  is a linear combination of
the square-free cyclic polynomials $C_p(\vx)$:
\[
  K(\vx) = \sum_{p \in \Bset} f_p(R) \, C_p(\vx),
\]
  where $\Bset$ % = \{0, 1, 12, 13, \ldots, 12\dots n\}$ is
  is the set of indices of all square-free cyclic polynomials,
  and $f_p(R)$ are polynomials of $R$.
  \label{thm:sqrfree}
\end{theorem}

\begin{proof}
Consider the following square-free reduction.
%
Given a cyclic polynomial $K(\vx)$,
  we recursively apply \refeq{logmaps} as
  $x_k^2 \rightarrow R - x_{k+1}$,
  until all squares or higher powers of $x_k$ are eliminated.
The process will not last indefinitely
  for each substitution reduces the degree in $x_k$ %(no matter which $k$)
  by one.
Since the original polynomial is cyclic,
  so is the reduced one, which is also square-free.
All terms that involve no $x_k$
  are collected to serve as the coefficient before $C_0(\vx) \equiv 1$.
%Since no square or higher powers of $x_k$
%  can survive the reduction,
%  the final cyclic polynomials are square-free;
The coefficients $f_p$ are polynomials of $R$,
  for $R$ is the only variable introduced by the substitutions.
\end{proof}

@@

\refthm{sqrfree}
  shows that any cyclic polynomial
  can be expanded
  as a combination of the square-free ones,
  which serve as a basis.
Below we show that at the onset and bifurcation points,
  the square-free cyclic polynomials $C_p(\vx)$ are themselves
  linearly connected by $\NB$ homogeneous equations.
The determinant of the matrix must vanish,
  and this condition yields the solution of the problem.

@@

We first observe that the derivative of $f^n$
is a cyclic polynomial:
%
\begin{equation}
  \Lambda(\vx)
   = \frac{d}{dx}f^{n}(x_1)
   = f'(x_n) \dots f'(x_1)
   = (-2)^n x_1 \dots x_n.
\label{eq:logder}
\end{equation}

Now, for any $p$, $\Lambda(\vx) \, C_p(\vx)$
  is also a cyclic polynomial,
since the product of two cyclic polynomials is cyclic too.
We can therefore expand it by \refthm{sqrfree} as
\begin{equation}
  \Lambda(\vx) \, C_p(\vx) = \sum_{q \in \Bset} T_{pq}(R) \, C_q(\vx),
\label{eq:xcp0}
\end{equation}
where $T_{pq}(R)$ is a polynomial of $R$,
and $p, q \in \Bset$.

By \refeq{der}, at the onset or bifurcation point,
  $\Lambda(\vx)$ is equal to a number
  $\lambda = +1$ or $-1$, respectively;
  so \refeq{xcp0}
  becomes a homogeneous linear equation of
  $C_p(\vx)$:
\begin{equation}
  \lambda \, C_p(\vx) = \sum_{q \in \Bset} T_{pq}(R) \, C_q(\vx),
\label{eq:xcp}
\end{equation}
%
%
%where $\lambda$ is $+1$ and $-1$ at the onset and bifurcation points, respectively;
or in matrix form,
\begin{equation}
  \big[ \lambda \, \vct I - \vct T(R) \big] \, \vct C = 0,
\tag{$\ref{eq:xcp}'$}
\end{equation}
%
%
where
$\vct I$ is the $\NB \times \NB$ identity matrix,
$\vct T(R) = \{T_{pq}(R)\}$ is an $\NB \times \NB$ matrix,
and
$\vct C = \{C_p(\vx)\}$ is an $\NB$-dimensional column vector.

Since a set of homogeneous linear equations
  has a non-trivial solution
  only if the determinant of the coefficient matrix
  is zero, we have
\begin{equation}
  A_n\RX \equiv \Big| \lambda \, \vct I - \vct T(R) \Big| = 0.
\label{eq:xr}
\end{equation}
Here, $A_n\RX$ is a polynomial of $R$ and $\lambda$,
  and the subscript $n$ is attached such that
  it can be used later with $A_d\RX$ of divisors $d$ of $n$.
\refeq{xr} is a necessary condition
  since $C_p(\vx)$ cannot vanish altogether;
and since it involves $R$ only,
  the polynomial expansion of the determinant
  gives the answer to our problem.
%Note that we did not solve the values $C_p$ in the process.

To summarize, we have
\begin{theorem}
  At the onset and bifurcation points,
  the square-free cyclic polynomials $C_p(\vx)$
  are linear related by \refeq{xcp}, with
  $\lambda$ being $+1$ and $-1$, respectively,
  and $T_{pq}(R)$ are the coefficients
  from the square-free reduction of $\Lambda(\vx) \, C_p(\vx)$
  with $\Lambda(\vx)$ specified by \refeq{logder}.
  Thus, $R$ at the two points are the roots of the
  polynomials $A_n(R, \lambda = \pm1)$
  obtained from the characteristic equation \refeq{xr}.
  \label{thm:main}
\end{theorem}

@@

We illustrate the above algorithm by cases of small $n$.
%
It is still helpful to have a mathematical software verify some steps
  (e.g., in computing the determinants and their factorization).

For $n = 1$, we have two square-free cyclic polynomials
  $C_0(\vx) = 1$ and $C_1(\vx) = x_1$;
  and $\Lambda(\vx) = -2 \, x_1$
\big[we shall drop ``$(\vx)$'' below for convenience\big].
Thus $\Lambda \, C_0 = -2 \, x_1 = -2 \, C_1$,
$\Lambda \, C_1 = -2 \, x_1^2 = -2 \, R + 2 x_1 = -2 \, R \, C_0 + 2 \, C_1$,
or
\[
  \lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right)
  =
  \left( \begin{array}{cc}
  0     & -2 \\
  -2R   &  2
  \end{array}\right)
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right),
\]
and \refeq{xr} reads
\begin{equation}
0=  \left| \begin{array}{cc}
  \lambda     & 2          \\
  2 R         & \lambda - 2
  \end{array} \right| = -4R - 2\lambda + \lambda^2,
  \tag{\ref{eq:xr}-1}
\label{eq:xr1}
\end{equation}
which is just the equation for a fixed point.
The fixed point begins at $\lambda = +1$ or $R_a = -1/4$,
  and becomes impossible when $\lambda = -1$, or $R_b = 3/4$.

For $n = 2$, the cyclic variables are
$C_0 = 1$,
$C_1 = x_1 + x_2$,
$C_{12} = x_1 x_2$,
and $\Lambda = 4 x_1 x_2$.
Thus,
$\Lambda \, C_0 = 4 C_{12}$,
$\Lambda \, C_1 = 4 \, {x_1}^2 \, x_2 + 4 \, x_1 \, {x_2}^2
  = 4 \, R \, (x_1 + x_2) - 4 ({x_1}^2 + {x_2}^2)
  = 4 \, R \, (x_1 + x_2) - 8 \, R + 4 \, (x_1 + x_2)
  = 4 \, (R + 1) \, C_1 - 8 \, R \, C_0$,
$\Lambda \, C_{12}
  = 4 \, (R-x_2)\, (R-x_1)
  = 4\, R^2  \, C_0 - 4 \, R \, C_1 + 4 \, C_{12}$,
or
\[
\lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1 \\
  C_{12}
  \end{array} \right)
 =
  \left( \begin{array}{ccc}
  0           & 0         & 4 \\
  -8R         & 4(R+1)    & 0 \\
  4R^2        & -4R       & 4
  \end{array} \right)
  \left( \begin{array}{c}
  C_0 \\
  C_1 \\
  C_{12}
  \end{array} \right),
\]
%
%
%
and \refeq{xr} reads
%
%
%
\begin{align}
0 &=
  \left| \begin{array}{ccc}
  \lambda     & 0               & -4 \\
  8R          & \lambda-4(R+1)  & 0 \\
  -4R^2       & 4R              & \lambda-4
  \end{array} \right| \nonumber\\
  &= (4R - 4 + \lambda) [(4 R-\lambda)^2 - 4 \lambda].
  \tag{\ref{eq:xr}-2}
\label{eq:xr2}
\end{align}
%
%
%
We only use the first factor
(the choice will be explained later, same for the following cases).
%
Setting it to zero yields $R = 1 - \lambda/4$;
$\lambda = +1$ gives the onset value $R_a = 3/4$ ($r_a = 3$)
while
$\lambda = -1$ gives the bifurcation value $R_b = 5/4$ ($r_b = 1+\sqrt 6$).
%
Note that the onset point of the only 2-cycle
  is located at $R = 3/4$,
  where the fixed point bifurcates \cite{strogatz}
  \big[compare \reffigs{cobweb}(a) and (b)\big].

For $n = 3$ \cite{saha, bechhoefer, gordon, burm, zhang},
we have
$C_0 = 1$,
$C_1 = x_1 + x_2 + x_3$,
$C_{12} = x_1 x_2 + x_2 x_3 + x_3 x_1$,
$C_{123} = x_1 x_2 x_3$,
and
$\Lambda = -8 x_1 x_2 x_3$.
%
The square-free reduction yields
%
%
%
\begin{align}
\lambda
\left( \begin{array}{c}
  C_0 \\
  C_1 \\
  C_{12} \\
  C_{123}
\end{array} \right)
=
\left( \begin{array}{cccc}
 0            & 0                 & 0               & -8 \\
 -24 R        & 8(R+1)            & -8R             & 0 \\
 24 R^2       & -8R(R+2)          & 8(R+1)          & 0 \\
 -8R^3        & 8R^2              & -8R             & 8
  \end{array} \right)
\left( \begin{array}{c}
  C_0 \\
  C_1 \\
  C_{12} \\
  C_{123}
\end{array} \right),
\end{align}
%
%
%
and \refeq{xr} reads:
\begin{align}
  0
&=\left| \begin{array}{cccc}
 \lambda      & 0                 & 0               & 8 \\
 24 R         & \lambda - 8(R+1)  & 8R              & 0 \\
 -24 R^2      & 8R(R+2)           & \lambda-8(R+1)  & 0 \\
 8R^3         & -8R^2             & 8R              & \lambda-8
  \end{array} \right| \nonumber \\
&= -\big[
  64R^3 - 128 R^2 - 8 (\lambda - 8) R - (\lambda -8)^2
  \big]
  \big(
  \lambda^2 - 8\lambda - 24 R\lambda - 64 R^3
  \big).
  \tag{\ref{eq:xr}-3}
\label{eq:xr3}
\end{align}
%
Using the first factor, we find at the onset point
  $\lambda = 1$,
  $\left(
  R-\frac74
  \right)
  \left(
  R^2-\frac14 R + \frac{7}{16}
  \right)=0$
  and its only real solution is $R_a = 7/4$ ($r_a = 1+\sqrt 8$).
%
At the bifurcation point $\lambda = -1$,
  the equation $R^3-2R^2 + \frac{9}{8} R -\frac{81}{64}=0$
  yields
  $R_b = \frac14
      \left(
      \frac{8}{3} +
        \sqrt[3]{\frac{1915}{54} - \frac52\sqrt{201}}
       +\sqrt[3]{\frac{1915}{54} + \frac52\sqrt{201}}
      \right)$,
whose corresponding $r = 1+\sqrt{1+4R}$
is identical to that in ref. \cite{gordon, burm}.
%

>>>>

In an $n$-cycle, $n$ is the smallest positive integer
  that allows $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
Thus, any $x_k$ in an $n$-cycle of $f$ must be a fixed point of $f^n$
%
\big[the converse is, however, untrue, for a fixed point of $f^n$
  can also be a fixed point of $f^d$ for a divisor $d$ of $dn$:
  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big],
%
%
and a cycle can be classified as stable or unstable
  according to the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_1) \cdots f'(x_n) \Big| \le 1,
\label{eq:der}
\end{equation}
The set of the $n$ points within the cycle $x_1. \ldots, x_n$
   is called the \emph{orbit}.
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz}.
%
%
%

@@

To simplify the calculation, we first change variables \cite{saha} as
%
%\begin{equation}
\[
    x^{\mathrm{new}} \equiv r(x^{\mathrm{old}} - 1/2),
    \quad R \equiv r(r-2)/4,
\]
%\label{eq:chvars}
%\end{equation}
%
and rewrite the map, in terms of $x^{\mathrm{new}}$, as
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv R - {x_k}^2.
  %\tag{$\ref{eq:logmap}'$}
\label{eq:logmaps}
\end{equation}
%
%\refeq{logmaps} simplifies \refeq{logmap} as $x_k$ now only occurs once.
%
We will solve the cycle boundaries as zeros of some polynomials of $R$,
and the corresponding polynomials for $r$ can be obtained by $R \rightarrow r(r-2)/4$.
%
%the numerical $r$ values can always be obtained from $r = 1 + \sqrt{1 + 4R}$.
%

@@

%Let us consider the first step of finding the fixed points of $f^n$.
At the first glance, the polynomials of the fixed points of $f^n$
  can be obtained by brute force.
We can solve
  \refeqs{logmaps} and express $x_1, \ldots, x_n$
  in terms of $R$,
  and then plug the solution into \req{der}.
The result would contain $R$ only,
  and is therefore the answer.
But since \refeqs{logmaps} are nonlinear,
  it is nontrivial to reduce the final equation of $R$
  into a polynomial one.
Nonetheless, on a computer, one can
  construct a (polynomial) Gr\"obner basis \cite{kk1}
  that gradually eliminates $x_k$ out of the $n+1$ equations.
As a special case, if $x_2, \dots, x_n$ are eliminated first,
  the final elimination of $x_1$ can be accomplished
  by setting the resultant of $f^n(x_1) = x_1$ and
  \refeq{der} or $\frac{d}{dx}f^n (x_1) = \pm 1$
  to zero \cite{burm},
  which furnishes the needed polynomial equation of $R$.
%
These approaches, however, do not
  exploit the cyclic structure of \refeqs{logmaps},
  and can thus be improved by the following alternative.

@@

A polynomial is cyclic if it is invariant
  under the cycling of variables
  $x_1 \rightarrow x_2, x_2 \rightarrow x_3,
  \ldots, x_n \rightarrow x_1$,
  e.g., $a(\vx) = x_1 x_2 + x_2 x_3 + x_3 x_4 + x_4 x_1$
  and $b(\vx) = x_1 x_3 + x_2 x_4$,
  where $\vx \equiv \{x_1, \ldots, x_n\}$ and $n = 4$.
Note that a cyclic polynomial is not necessarily symmetric
  (i.e., invariant under the exchange of any two $x_k$ and $x_{j}$),
  e.g., $a(\vx)$ and $b(\vx)$ are not symmetric,
  but $a(\vx) + b(\vx)$ is.

@@

A cyclic polynomial is \emph{square-free}
  if it has no square or higher powers of $x_k$,
e.g., $C_{12}(\vx) = x_1 x_2 + \cdots$ is square-free,
  $C_{112}(\vx) = {x_1}^2 x_2 + \cdots$ is not,
  see \reftab{sqrfreepoly} for more examples.
The label $p$ of a square-free polynomial
  has no repeated index.
We denote the set of all square-free $p$ by
$\Pset = \{0, 1, 12, 13, \ldots, 123, 124, \ldots, 12\dots n\}$,
and its size $|\Pset|$ equals the number $\NB$ of
  square-free cyclic polynomials.

@@

The set of square-free cyclic polynomials
  furnishes a basis for expanding cyclic polynomials:

\begin{theorem}
  For the logistic map \refeq{logmaps},
  any cyclic polynomial $K(\vx)$
  formed by the $n$-cycle points
  $\vct x = \{x_1, \ldots, x_n\}$
  is a linear combination of
the square-free cyclic polynomials $C_p(\vx)$:
\[
  K(\vx) = \sum_{p \in \Pset} f_p(R) \, C_p(\vx),
\]
  where $\Pset$ % = \{0, 1, 12, 13, \ldots, 12\dots n\}$ is
  is the set of indices of all square-free cyclic polynomials,
  and $f_p(R)$ are polynomials of $R$.
  \label{thm:sqrfree}
\end{theorem}

\begin{proof}
Consider the following square-free reduction.
%
Given a cyclic polynomial $K(\vx)$,
  we recursively apply \refeq{logmaps} as
  $x_k^2 \rightarrow R - x_{k+1}$,
  until all squares or higher powers of $x_k$ are eliminated.
The process will not last indefinitely
  for each substitution reduces the degree in $x_k$ %(no matter which $k$)
  by one.
Since the original polynomial is cyclic,
  so is the final square-free polynomial.
Terms involving no $x_k$
  are collected to serve as the coefficient before $C_0(\vx) \equiv 1$.
%Since no square or higher powers of $x_k$
%  can survive the reduction,
%  the final cyclic polynomials are square-free;
The coefficients $f_p$ are polynomials of $R$,
  for $R$ is the only variable introduced by the substitutions.
\end{proof}

@@

\refthm{sqrfree}
  shows that any cyclic polynomial can be expanded
  as a combination of the square-free ones.
Below we show that at the onset and bifurcation points,
  the square-free cyclic polynomials $C_p(\vx)$ are themselves
  linearly connected by $\NB$ homogeneous equations.
The determinant of the matrix must vanish,
  and this condition yields the solution of the problem.

@@

We first observe that the derivative of $f^n$
is a cyclic polynomial:
%
%
\begin{equation}
  \Lambda(\vx)
   = \frac{d}{dx}f^{n}(x_1)
   = f'(x_n) \dots f'(x_1)
   = (-2)^n x_1 \dots x_n.
\label{eq:logder}
\end{equation}
%
%
Since $C_p(\vx)$ is also cyclic,
  so is the product $\Lambda(\vx) \, C_p(\vx)$,
  which can be expanded by \refthm{sqrfree} as
\begin{equation}
  \Lambda(\vx) \, C_p(\vx) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx),
\label{eq:xcp0}
\end{equation}
where $T_{pq}(R)$ is a polynomial of $R$,
and $p, q \in \Pset$.

By \refeq{der}, at the onset or bifurcation point,
  $\Lambda(\vx)$ is equal to a number
  $\lambda = +1$ or $-1$, respectively;
  so \refeq{xcp0}
  becomes a homogeneous linear equation of
  $C_p(\vx)$:
\begin{equation}
  \lambda \, C_p(\vx) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx),
\label{eq:xcp}
\end{equation}
%
%
%where $\lambda$ is $+1$ and $-1$ at the onset and bifurcation points, respectively;
or in matrix form,
\begin{equation}
  \big[ \lambda \, \vct I - \vct T(R) \big] \, \vct C = 0,
\tag{$\ref{eq:xcp}'$}
\end{equation}
%
%
where
$\vct I$ is the $\NB \times \NB$ identity matrix,
$\vct T(R) = \{T_{pq}(R)\}$ is an $\NB \times \NB$ matrix,
and
$\vct C = \{C_p(\vx)\}$ is an $\NB$-dimensional column vector.

Since a set of homogeneous linear equations
  has a non-trivial solution
  only if the determinant of the coefficient matrix
  is zero, we have
\begin{equation}
  A_n\Rlam \equiv \Big| \lambda \, \vct I - \vct T(R) \Big| = 0.
\label{eq:xr}
\end{equation}
Here, $A_n\Rlam$ is a polynomial of $R$ and $\lambda$,
  and the subscript $n$ is attached such that
  it can be used later with $A_d\Rlam$ of divisors $d$ of $n$.
\refeq{xr} is a necessary condition
  since $C_p(\vx)$ cannot vanish altogether;
and since it involves $R$ only,
  the polynomial expansion of the determinant
  gives the answer to our problem.
%Note that we did not solve the values $C_p$ in the process.

To summarize, we have
\begin{theorem}
  At the onset and bifurcation points,
  the square-free cyclic polynomials $C_p(\vx)$
  of the orbit points $\vx = \{x_k\}$
  are linear related by % \refeq{xcp},
  $\lambda \, C_p(\vx) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx)$,
  with $\lambda$ being $+1$ and $-1$, respectively,
  and $T_{pq}(R)$ being the coefficients
  from the square-free reduction of the cyclic polynomial
  $C_p(\vx) \, \prod_k f'(x_k)$. %[\refeq{logder}].
  Thus, the fixed point of $f^n$
  is marginally stable when $R$ satisfies
  characteristic equation $\big| \lambda \, \vct I - \vct T(R) \big| = 0$.
  \label{thm:main}
\end{theorem}

@@

We illustrate the above algorithm by several small-$n$ cases.
%
It is still helpful to have a mathematical software
  in order to verify certain steps
  (e.g., in computing the determinants and their factorization).

For $n = 1$, we have two square-free cyclic polynomials
  $C_0(\vx) = 1$ and $C_1(\vx) = x_1$;
  and $\Lambda(\vx) = -2 \, x_1$
\big[we shall drop ``$(\vx)$'' below for convenience\big].
Thus $\Lambda \, C_0 = -2 \, x_1 = -2 \, C_1$,
$\Lambda \, C_1 = -2 \, x_1^2 = -2 \, R + 2 x_1 = -2 \, R \, C_0 + 2 \, C_1$,
or
\[
  \lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right)
  =
  \left( \begin{array}{cc}
  0     & -2 \\
  -2R   &  2
  \end{array}\right)
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right),
\]
and \refeq{xr} reads
\begin{equation}
0=  \left| \begin{array}{cc}
  \lambda     & 2          \\
  2 R         & \lambda - 2
  \end{array} \right| = -4R - 2\lambda + \lambda^2,
  \tag{\ref{eq:xr}-1}
\label{eq:xr1}
\end{equation}
which is just the equation for a fixed point.
The fixed point begins at $\lambda = +1$ or $R_a = -1/4$,
  and becomes unstable when $\lambda = -1$, or $R_b = 3/4$.

For $n = 2$, the cyclic polynomials are
$C_0 = 1$,
$C_1 = x_1 + x_2$,
$C_{12} = x_1 x_2$,
and $\Lambda = 4 x_1 x_2$.
Thus,
$\Lambda \, C_0 = 4 C_{12}$,
$\Lambda \, C_1 = 4 \, {x_1}^2 \, x_2 + 4 \, x_1 \, {x_2}^2
  = 4 \, R \, (x_1 + x_2) - 4 ({x_1}^2 + {x_2}^2)
  = 4 \, R \, (x_1 + x_2) - 8 \, R + 4 \, (x_1 + x_2)
  = 4 \, (R + 1) \, C_1 - 8 \, R \, C_0$,
$\Lambda \, C_{12}
  = 4 \, (R-x_2)\, (R-x_1)
  = 4\, R^2  \, C_0 - 4 \, R \, C_1 + 4 \, C_{12}$,
or
\[
\lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1 \\
  C_{12}
  \end{array} \right)
 =
  \left( \begin{array}{ccc}
  0           & 0         & 4 \\
  -8R         & 4(R+1)    & 0 \\
  4R^2        & -4R       & 4
  \end{array} \right)
  \left( \begin{array}{c}
  C_0 \\
  C_1 \\
  C_{12}
  \end{array} \right),
\]
%
%
%
and \refeq{xr} reads
%
%
%
\begin{align}
0 &=
  \left| \begin{array}{ccc}
  \lambda     & 0               & -4 \\
  8R          & \lambda-4(R+1)  & 0 \\
  -4R^2       & 4R              & \lambda-4
  \end{array} \right| \nonumber\\
  &= (4R - 4 + \lambda) [(4 R-\lambda)^2 - 4 \lambda].
  \tag{\ref{eq:xr}-2}
\label{eq:xr2}
\end{align}
%
%
%
We only use the first factor
(the choice will be explained later, same for the following cases).
%
Setting it to zero yields $R = 1 - \lambda/4$;
$\lambda = +1$ gives the onset value $R_a = 3/4$ ($r_a = 3$)
while
$\lambda = -1$ gives the bifurcation value $R_b = 5/4$ ($r_b = 1+\sqrt 6$).
%
Note that the onset point of the only 2-cycle
  is located at $R = 3/4$,
  where the fixed point bifurcates \cite{strogatz}
  \big[compare \reffigs{cobweb}(a) and (b)\big].

For $n = 3$ \cite{saha, bechhoefer, gordon, burm, zhang},
we have
$C_0 = 1$,
$C_1 = x_1 + x_2 + x_3$,
$C_{12} = x_1 x_2 + x_2 x_3 + x_3 x_1$,
$C_{123} = x_1 x_2 x_3$,
and
$\Lambda = -8 x_1 x_2 x_3$.
%
The square-free reduction yields
%
%
%
\begin{align*}
\lambda
\left( \begin{array}{c}
  C_0 \\
  C_1 \\
  C_{12} \\
  C_{123}
\end{array} \right)
=
\left( \begin{array}{cccc}
 0            & 0                 & 0               & -8 \\
 -24 R        & 8(R+1)            & -8R             & 0 \\
 24 R^2       & -8R(R+2)          & 8(R+1)          & 0 \\
 -8R^3        & 8R^2              & -8R             & 8
  \end{array} \right)
\left( \begin{array}{c}
  C_0 \\
  C_1 \\
  C_{12} \\
  C_{123}
\end{array} \right),
\end{align*}
%
%
%
and \refeq{xr} reads:
\begin{align}
  0
&=\left| \begin{array}{cccc}
 \lambda      & 0                 & 0               & 8 \\
 24 R         & \lambda - 8(R+1)  & 8R              & 0 \\
 -24 R^2      & 8R(R+2)           & \lambda-8(R+1)  & 0 \\
 8R^3         & -8R^2             & 8R              & \lambda-8
  \end{array} \right| \nonumber \\
&= -\big[
  64R^3 - 128 R^2 - 8 (\lambda - 8) R - (\lambda -8)^2
  \big]
  \big(
  \lambda^2 - 8\lambda - 24 R\lambda - 64 R^3
  \big).
  \tag{\ref{eq:xr}-3}
\label{eq:xr3}
\end{align}
%
Using the first factor, we find at the onset point
  $\lambda = 1$,
  $\left(
  R-\frac74
  \right)
  \left(
  R^2-\frac14 R + \frac{7}{16}
  \right)=0$
  and its only real solution is $R_a = 7/4$ ($r_a = 1+\sqrt 8$).
%
At the bifurcation point $\lambda = -1$,
  the equation $R^3-2R^2 + \frac{9}{8} R -\frac{81}{64}=0$
  yields
  $R_b = \frac14
      \left(
      \frac{8}{3} +
        \sqrt[3]{\frac{1915}{54} - \frac52\sqrt{201}}
       +\sqrt[3]{\frac{1915}{54} + \frac52\sqrt{201}}
      \right)$,
whose corresponding $r = 1+\sqrt{1+4R}$
is identical to that in ref. \cite{gordon, burm}.
%

For $n = 4$ \cite{stephenson, bailey1, kk1, bailey2, lewis},
the cyclic polynomials are
$C_0 = 1$,
$C_1 = x_1 + x_2 + x_3 + x_4$,
$C_{12} = x_1 x_2 + x_2 x_3 + x_3 x_4 + x_4 x_1$,
$C_{13} = x_1 x_3 + x_2 x_4$,
$C_{123} = x_1 x_2 x_3 + x_2 x_3 x_4 + x_3 x_4 x_1 + x_4 x_1 x_2$,
$C_{1234} = x_1 x_2 x_3 x_4$;
and
$\Lambda = 16 x_1 x_2 x_3 x_4$.
%
\refeq{xr} reads
%
\begin{align}
0 & = \left|
\begin{array}{cccccc}
 \lambda  & 0               & 0                 & 0     & 0     & -16 \\
 64 R     & \lambda-16(R+1) & 16 R              & 0     & -16R  & 0 \\
 -64 R^2  & 16R(R+2)        & \lambda-16(R^2+1) & -32 R & 16R   & 0 \\
 -32 R^2  & 16R(R+1)        & -16R              & \lambda-16(R^2+1)  & 0 & 0 \\
 64 R^3   & -16R^2(R+3)     & 16R(R+2)        & 32 R (R+1)       & \lambda-16(R+1) & 0 \\
 -16R^4   & 16R^3           & - 16R^2           & -16R^2            & 16R & \lambda - 16
\end{array}
\right | \nonumber  \\
& =
    \big[
    4096 R^6-12288 R^5 + 256 (\lambda + 48) (R^4- R^3)
      -16 (\lambda + 32)(\lambda - 16) R^2
      -(\lambda - 16)^3 \big]  \nonumber \\
& \quad
    \big[ 16 (R-1)^2 - \lambda \big]
 \, \big[(16 R^2 + \lambda)^2 - 16 (2R+1)^2 \lambda \big],
  \tag{\ref{eq:xr}-4}
\label{eq:xr4}
\end{align}
%
%
%
From the first factor, we have
$ (
    4R - 5
  )
\, \big[
  (4R + 1)^2 + 4
  \big]
\, \big[
  (4R - 3)^3 - 108
  \big] =0$
at the onset point $\lambda = 1$.
%
%
%
It has two real roots:
$R_a' = 5/4$
\big($r_a' = 1+\sqrt{6} \approx 3.4495$\big)
for the cycle from period-doubling the 2-cycle
  \big[compare \reffigs{cobweb}(b) and (d)\big],
and
$R_a = (3+\sqrt[3]{108})/4$
\big($r_a = 1+\sqrt{4+\sqrt[3]{108}} \approx 3.9601$\big)
for an original cycle \big[\reffig{cobweb}(e)\big].
%
At the bifurcation point, $\lambda = -1$, and
$4096 R^6 - 12288 R^5 + 12032 (R^4 - R^3)
  + 8432 R^2 + 4913 = 0$,
which upon $R \rightarrow r(r-2)/4$ yields the same polynomial
obtained previously
\cite{bailey1, kk1, bailey2, lewis}.
%
%
%
The only two positive roots $r_b \approx 3.9608$
and $r_b' \approx 3.5441$
correspond to $r_a$ and $r_a'$ respectively.
%
As a verification, the polynomials are
alternatively derived in \refapd{per4}.
